### 2014-10-21 Projects/NFRL_Emissions/Report/ECS_Testing.tex
Reports: Additional minor edits; update template
NFRL ECS: update report, formatting in bib file, update figure

\headerC{
{
\flushright{
Craig G. Weinschenk \\
Matt Bundy \\

@@

\begingroup
\hypersetup{urlcolor=black}
This publication is available free of charge from:
\href{http://dx.doi.org/10.6028/NIST.TN.XXXX}{http://dx.doi.org/10.6028/NIST.TN.XXXX} \\
\endgroup

@@

\begin{tabbing}
\hspace{1.5in} \= \\
FDS \> Fire Dynamics Simulator \\
HRR \> Heat Release Rate \\
NIST \> National Institute of Standards and Technology \\
\end{tabbing}

@@

\chapter{Introduction}
\pagenumbering{arabic}
\setcounter{page}{1}
The National Fire Research Laboratory (NFRL) is a fire test facility operated by the Engineering Laboratory at the National Institute of Standards and Technology (NIST). The NFRL is undergoing an expansion in order to provide additional research capabilities: test the performance of full-scale structures subjected to realistic fires and structural loading under controlled laboratory conditions, develop an experimental database on the performance of large-scale structural connections, components, sub-assemblies, and systems under realistic fire and loading, validate physics-based models to predict fire resistance performance of structures and enable performance-based standards for fire resistance design of structures and foster innovations in design and construction. To complete construction on the NFRL expansion, the updated emission control system needs to be tested to ensure the design specifications are achieved. This document will describe the source fires, appropriate lab safety, and test procedures. The objective of the fire tests is to use specifically designed source fires to test the emission control systems of the newly constructed fire research facility addition. The experiments are designed to ensure the system can handle elevated temperatures (gas burner), toxic gases (wood cribs with PVC), and particulates (group A commodities).

\chapter{Test Fires}
\label{test_fire}
The commissioning fires will take place in the pre-expansion laboratory space under a 9 m by 12 m oxygen-consumption calorimetry hood. The fuel packages will be centered under the hood, which is 6.4 m above the test floor. A flexible curtain, in place to contain combustion products, hangs 1.5 m below the lower edge of hood. Fig.~(\ref{fig:NFRL_fire}) shows a 2 MW natural gas fire under the hood during an experiment prior to the lab expansion. Note, that in the current configuration the wall to the left of the burner has been replaced by a roll-up door. The roll-up door provides access to the new testing space.

\begin{figure}
\includegraphics[width=\textwidth]{../Figures/NFRL_Fire}
\caption {Natural gas test burn in pre-construction NFRL.}
\label{fig:NFRL_fire}
\end{figure}

Based on the general experimental configuration shown in Fig.~(\ref{fig:NFRL_fire}) and \ref{app:nfrl_images}, Table ~\ref{tab:fires} outlines the fire tests to be conducted. The duplicity of tests is that similar to the existing system, the new ECS features two trains. Therefore, the scrubbing system in each train needs to be tested independently.

\begin{table}
\centering
\caption{Design fires for ECS commissioning tests}
\label{tab:fires}
\begin{tabular}{ccccc}
\toprule[1.5pt]
Test ID & Fuel & Peak HRR (MW) & Burn Time (min) & Test Goal  \\
\midrule
1-3  & Natural Gas  & 2-8   & 60  & Temperature  \\
4  & Wood w/ PVC  & $\approx$ 1.5-2 & $\approx$ 50  & Acid, Train 3 \\
5  & Group A Comm. & $\approx$ 3 & $\approx$ 20  & Soot, Train 3 \\
6  & Wood w/ PVC  & $\approx$ 1.5-2 & $\approx$ 50  & Acid, Train 4 \\
7  & Group A Comm. & $\approx$ 3 & $\approx$ 20  & Soot, Train 4 \\
\bottomrule[1.25pt]
\end{tabular}\par
\end{table}

\section{Test 1}
\label{test1}
A natural gas burner is a very clean and well controlled source fire (generating CO$_2$ and H$_2$O products). The objective of this test is to challenge the system for hot gases using a controllable and well quantified fire. This is not designed to test scrubbing capabilities, but to prove the system can achieve the required temperature set points. It is critical to ensure the scrubber is working properly before pollutants are introduced. The natural gas burner is 1.61 m by 1.19 m by 0.88 m and is capable of delivering a fire up to 8 MW (cf. Fig.~(\ref{fig:nat_gas_burn})).

@@

The required mass flow of natural gas ($\dot{m}_{NG}$) needed to produce a desired fire size ($HRR_{NG}$) is
\begin{equation}
\dot{m}_{\mathrm{NG}} = HRR_{NG}/h_{c_{NG}}
\end{equation}
where $h_{c_{NG}}$ is the heat of combustion of natural gas. For complete combustion, the heat of combustion of natural gas is 50203 kJ/kg. Based on calculations included in \ref{app:gas_burn}, the required mass flow rate of natural gas for range of fire sizes is shown in Table~\ref{tab:ng_flow}. While the peak deliverable HRR is 8 MW, radiation from the fire to the surrounding environment at this magnitude is too significant for a 60 minute commissioning test. To test each train, a 4 MW fire will be used. To test that both trains can properly work together, an 8 MW fire will be used for a short duration, approximately 5 minutes.

@@

\section{Test 2}
\label{test2}
To challenge the acid scrubbing capability of the system, pinewood cribs loaded with polyvinyl chloride (PVC) are used. This package is designed to deliver a controlled source of hydrochloric acid (HCl) gas that is close to the design capacity of the emission control system. Four pinewood cribs measuring 0.56 m by 0.56 m by 0.38 m will be used (cf. Fig.~(\ref{fig:wood_crib})). Each crib consists of 10 layers composed of 0.56 m by 0.04 m by 0.04 m pinewood sticks spaced 0.02 m apart. Eight 0.45 kg PVC sticks are distributed throughout each crib (cf. Fig.~(\ref{fig:wood_crib})) which brings the total mass of each crib to be 34 kg. The moisture content of the cribs should be between 6-8 \% using a commercial moisture detector. This fuel package is based on the wood crib tests from the 2001 commissioning tests.

@@

To determine the HCL yield of PVC, we need to determine a balanced chemical reaction. First, is to assume that the test PVC can be represented by the chemical formula, $\C_2\H_3\Cl$ and that all of the Cl in PVC goes to HCl. Then, based on work by Tewarson \cite{SFPE:Tewarson}, yields for soot and CO are found ($y_{\mathrm{C}}$= 0.172, $y_{\mathrm{CO}}$= 0.063) and a balanced chemical reaction is created:

\begin{multline}
\C_2\H_3\Cl \,+\, 7.29(0.21 \O_2 \,+\, 0.79 \N_2) \\
\rightarrow 9.76(0.102\HCl \,+\, 0.103\H_2\O \,+\, 0.014 \C\O \,+\, 0.098 \C\O_2 \,+\, 0.092 \C \,+\, 0.59 \N_2)
\label{eq:hcl_reac}
\end{multline}
Based on Eq.~(\ref{eq:hcl_reac}), there is a 58 \% mass basis theoretical yield of HCl (\ref{app:crib}).  Experimental data based on burning PVC over a wide range of test conditions and using a wide range of measurement techniques puts the HCl yield between 10 and 56 \% \cite{Leeds:PVC}. Using data for a similarly scaled and ventilated test, the estimated HCl yield is 32 \% \cite{Persson1998}. Since each crib contains 3.63 kg of PVC and the test calls for 4 cribs, the expected yield of HCl at the source fire is 4.65 kg (10.2 lb). Based on an estimated 60-70 \% loss of HCl to gas phase conversion and reaction through the duct work, the estimated mass of HCL upstream of the scrubber is 1.4-1.9 kg (3-4 lb).

The 4 cribs will be setup in two tiers. The first tier will be three cribs arranged in a triangle. The second tier will use the forth crib centered over the void created by the triangle. Each of the lower tier cribs will be raised 0.08 m off the ground to allow for a pan of heptane to slide underneath. The test configuration is shown in Fig.~(\ref{fig:test_crib}).The cribs are ignited using 500 mL of heptane poured into a round steel pan. The heptane represents about 1.5 \% of the initial mass of the crib, and burns out after about 2.5 to 3 minutes. The forth crib will ignite as a result of flame spread from the other three cribs.

@@

\section{Test 3}
\label{test3}
The third commissioning test is designed to test the system's capability to handle particulates, specifically soot. Standard Group A commodity boxes will be used as the fuel source. Each box, is a cube 0.53 m on a side and is filled with a plastic commodity. The plastic commodity inside the box consists of rigid crystalline polystyrene cups (empty, 0.47 L (16 fl oz) size) packaged in compartmented, single-wall, corrugated cardboard box. The cups are arranged open end down in five layers, 25 per layer for a total of 125 per box (cf. Fig.~(\ref{fig:open_box})).

@@

In these tests, the columns of boxes are spaced 0.15 m apart (cf. Fig.~(\ref{fig:dim_box})). It is important to note that both orientation and spacing of these boxes can have a significant impact on the burning characteristics \cite{Madrzykowski2013}.

@@

The two columns of boxes will be ignited using shredded paper placed in the gap between the two columns of boxes. The paper will be ignited by a 1.83 m (6 ft) propane wand. As the cardboard burns away, the columns will lose structural integrity and the fire will transition into a polystyrene pool fire. The ignition point is set at the center line so that as the columns begin to collapse, they will collapse toward the center as that cardboard burned away first. The experiment will take place on top of sheets of cement board (such as USG Durock). This is done to contain the polystyrene melt in the later portion of the fire test. Fig.~(\ref{fig:comm_burn}) shows two columns of Group A commodity shortly after ignition and when the boxes are fully involved.

@@

\appendix
\chapter{Schematics of the National Fire Research Lab}
\label{app:nfrl_images}
\begin{figure}[!ht]
\includegraphics[width=\textwidth]{../Figures/plan_view2}
\caption {Plan view of the NFRL indicating location of ECS test fires.}
\label{fig:NFRL_plan}
\end{figure}

>>>>

\headerC{
{
\flushright{
Craig G. Weinschenk \\
Matthew Bundy \\

@@

\vspace{0.2in}

\begingroup
{\bf This publication is available free of charge from:}
\hypersetup{urlcolor=black}
\href{http://dx.doi.org/10.6028/NIST.TN.XXXX}{\bf http://dx.doi.org/10.6028/NIST.TN.XXXX} \\
\endgroup

@@

\begin{tabbing}
\hspace{1.5in} \= \\
ECS \> Emisssion Control System \\
HRR \> Heat Release Rate \\
NFRL \> National Fire Research Laboratory \\
NIST \> National Institute of Standards and Technology \\
PVC \> Polyvinyl Chlorine \\
\end{tabbing}

@@

\chapter{Introduction}
\pagenumbering{arabic}
\setcounter{page}{1}
The National Fire Research Laboratory (NFRL) is a fire test facility operated by the Engineering Laboratory at the National Institute of Standards and Technology (NIST). The NFRL underwent an expansion in order to provide additional research capabilities: test the performance of full-scale structures subjected to realistic fires and structural loading under controlled laboratory conditions, develop an experimental database on the performance of large-scale structural connections, components, sub-assemblies, and systems under realistic fire and loading, validate physics-based models to predict fire resistance performance of structures and enable performance-based standards for fire resistance design of structures and foster innovations in design and construction. To commission the laboratory space of the NFRL expansion, the updated emission control system needed to be tested to ensure the design specifications are achieved. This document will describe the source fires, appropriate lab safety, and test procedures. The objective of the fire tests was to use specifically designed source fires to test the emission control systems to ensure the system could handle elevated temperatures, toxic gases, and particulates.

\chapter{Test Fires}
\label{test_fire}
The commissioning fires took place in the pre-expansion laboratory space under the 9 m by 12 m oxygen-consumption calorimetry hood (cf Fig.~\ref{fig:NFRL_plan}). The dark shaded area in Fig.~\ref{fig:NFRL_plan} represents the laboratory expansion. The fuel packages were centered under the hood, which is 6.4 m above the test floor. A flexible curtain, in place to contain combustion products, hangs 1.5 m below the lower edge of hood. Fig.~(\ref{fig:NFRL_fire}) shows a 2 MW natural gas fire under the calorimetry hood during the initial commission testing. Note, that the the roll-up door behind the burner provides protection to the additional lab space when closed and access to the expansion space when opened.

\begin{figure}[!ht]
\centering
\includegraphics[width=.8\textwidth]{../Figures/plan_view2}
\caption {Plan view of the NFRL indicating location of ECS test fires.}
\label{fig:NFRL_plan}
\end{figure}

\begin{figure}[!ht]
\centering
\includegraphics[width=.6\textwidth]{../Figures/NFRL_Fire}
\caption {Natural gas test burn in pre-expansion side of NFRL.}
\label{fig:NFRL_fire}
\end{figure}

Overall, three different fuel packages were used to test the system's cabilitity to independently handle elevated temperatures, scrub acid gas, and trap particulates. Table ~\ref{tab:fires} shows the fire tests to be conducted. Two tests were conducted with each fuel package to ensure that each train within the new ECS was independently tested. For the gas burner tests, three tests were conducted; one for each of the the trains and one with both trains running together to test heat release rates larger than 5~MW.

\begin{table}
\centering
\caption{Design fires for ECS commissioning tests}
\label{tab:fires}
\begin{tabular}{lccl}
\toprule[1.5pt]
Fuel & Peak HRR (MW) & Burn Time (min) & Test Goal  \\
\midrule
Natural Gas   & 2-8             & 60            & Temperature  \\
Wood w/ PVC   & $\approx$ 1.5-2 & $\approx$ 50  & Acid Gases \\
Group A Comm. & $\approx$ 3     & $\approx$ 20  & Soot \\
\bottomrule[1.25pt]
\end{tabular}\par
\end{table}

\section{Test 1}
\label{test1}
A natural gas burner is a clean burning and well controlled source fire (predominately generating CO$_2$ and H$_2$O products). The objective of this test was to challenge the system for hot gases using a controllable and quantifiable fire. This was not designed to test scrubbing capabilities, but to prove the system can achieve the required temperature set points without failure. It was critical to ensure the scrubber is functional before pollutants are introduced to test its performance. The natural gas burner used was 1.61 m by 1.19 m by 0.88 m and was capable of delivering a fire up to 8~MW (cf. Fig.~(\ref{fig:nat_gas_burn})).

@@

The required mass flow of natural gas ($\dot{m}_{NG}$) needed to produce a desired fire size ($HRR_{NG}$) is
\begin{equation}\label{eq:mass_ng}
\dot{m}_{\mathrm{NG}} = HRR_{NG}/h_{c_{NG}}
\end{equation}
where $h_{c_{NG}}$ is the heat of combustion of natural gas. For complete combustion, the heat of combustion of natural gas is 50203~kJ/kg. Using Eq.~(\ref{eq:mass_ng}), the required mass flow rate of natural gas for range of fire sizes is shown in Table~\ref{tab:ng_flow}. While the peak deliverable HRR is 8~MW, radiation from the fire that size to the surrounding environment would be too significant for a 60 min. commissioning test without add additional fire protection to the structure. To test each train, a fire with a peak 4~MW HRR was used. To test that both trains can properly work together, an 8 MW fire was used for duration of less than 5~min.

@@

\section{Test 2}
\label{test2}
To challenge the acid scrubbing capability of the system, pinewood cribs loaded with polyvinyl chloride (PVC) were used. This package was designed to deliver a controlled source of hydrochloric acid (HCl) gas that was close to the design capacity of the emission control system. Four pinewood cribs measuring 0.56~m by 0.56~m by 0.38~m were be used (cf. Fig.~\ref{fig:wood_crib}). Each crib consisted of 10 layers composed of 0.56~m by 0.04~m by 0.04~m pinewood sticks spaced 0.02~m apart. Eight 0.45~kg PVC sticks were distributed throughout each crib (cf. Fig.~\ref{fig:wood_crib}) which brought the total mass of each crib to be 34~kg. The moisture content of the cribs was be between 6-8 \%, (measured using a commercial moisture detector). This fuel package was based on the prior wood crib tests used in the inital commisioning of the building in 2001.

@@

To determine the HCL yield of PVC, the first step was to calculate a balanced chemical reaction. PVC can be represented by the chemical formula, $\C_2\H_3\Cl$. If it assumed that all of the Cl in PVC goes to HCl, then, based on work by Tewarson \cite{SFPE:Tewarson}, yields for soot and CO are found: ($y_{\mathrm{C}}$= 0.172, $y_{\mathrm{CO}}$= 0.063). A balanced chemical reaction is created:

\begin{multline}
\C_2\H_3\Cl \,+\, 7.29(0.21 \O_2 \,+\, 0.79 \N_2) \\
\rightarrow 9.76(0.102\HCl \,+\, 0.103\H_2\O \,+\, 0.014 \C\O \,+\, 0.098 \C\O_2 \,+\, 0.092 \C \,+\, 0.59 \N_2)
\label{eq:hcl_reac}
\end{multline}
Based on Eq.~(\ref{eq:hcl_reac}), there is a 58~\% mass basis theoretical yield of HCl.  Experimental data based on burning PVC over a wide range of test conditions and using a wide range of measurement techniques puts the HCl yield between 10 and 56~\% \cite{Leeds:PVC}. Using data for a similarly scaled and ventilated test, the estimated HCl yield is 32~\% \cite{Persson1998}. Since each crib contains 3.63~kg (8~lb) of PVC and the test calls for 4 cribs, the expected yield of HCl at the source fire is 4.65~kg (10.2~lb). {\em Based on an estimated 60-70 \% loss of HCl to gas phase conversion and reaction through the duct work, the estimated mass of HCL upstream of the scrubber is 1.4-1.9~kg (3-4~lb).}

The 4 cribs will be setup in two tiers. The first tier will be three cribs arranged in a triangle. The second tier will use the forth crib centered over the void created by the triangle. Each of the lower tier cribs will be raised 0.08~m off the ground to allow for a pan of heptane to slide underneath. The test configuration is shown in Fig.~\ref{fig:test_crib}.The cribs are ignited using 500~mL of heptane poured into a round steel pan. The heptane represents about 1.5~\% of the initial mass of the crib, and burns out after about 2.5 to 3~min. The forth crib will ignite as a result of flame spread from the other three cribs.

@@

\section{Test 3}
\label{test3}
The third commissioning test is designed to test the system's capability to handle particulates, specifically soot. Standard Group A commodity boxes will be used as the fuel source. Each box, is a cube 0.53 m on a side and is filled with a plastic commodity. The plastic commodity inside the box consists of rigid crystalline polystyrene cups (empty, 0.47 L (16 fl oz) size) packaged in compartmented, single-wall, corrugated cardboard box. The cups are arranged open end down in five layers, 25 per layer for a total of 125 per box (cf. Fig.~\ref{fig:open_box}).

@@

In these tests, the columns of boxes are spaced 0.15 m apart (cf. Fig.~\ref{fig:dim_box}). It is important to note that both orientation and spacing of these boxes can have a significant impact on the burning characteristics \cite{Madrzykowski2013}.

@@

The two columns of boxes will be ignited using shredded paper placed in the gap between the two columns of boxes. The paper will be ignited by a 1.83 m (6 ft) propane wand. As the cardboard burns away, the columns will lose structural integrity and the fire will transition into a polystyrene pool fire. The ignition point is set at the center line so that as the columns begin to collapse, they will collapse toward the center as that cardboard burned away first. The experiment will take place on top of sheets of cement board (such as USG Durock). This is done to contain the polystyrene melt in the later portion of the fire test. Fig.~\ref{fig:comm_burn} shows two columns of Group A commodity shortly after ignition and when the boxes are fully involved.

@@

\appendix
\chapter{Schematics of the National Fire Research Lab}
\label{app:nfrl_images}
